---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file carries out data analysis and visualization for a paper on documentation practices of R packages.

First, load in the relevant data and packages.

```{r load-data, warning = FALSE, message = FALSE}
# load in packages
library(tidyverse)
library(infer)

# load in data
packages <- read_csv("data/packages.csv")
functions <- read_csv("data/functions.csv")
load("data/packages_summary.Rda")
```

Taking a glimpse...

```{r glimpse}
glimpse(packages)

glimpse(functions)
```

First, we'll join these data to add on package camp and mutate some new columns that will be useful for analysis and visualization.

```{r join-mutate}
# add on tidy/not-tidy to the functions dataframe
functions <- packages %>%
  select(package, camp, is_tidy) %>%
  left_join(functions)

functions <- functions %>%
  mutate(n_char_per_function = case_when(
           n_functions != 0 & (!is.na(n_functions)) ~ n_char/n_functions,
           TRUE ~ NA_real_),
         examples_per_function = case_when(
           n_functions != 0 & (!is.na(n_functions)) ~ n_examples /n_functions,
           TRUE ~ NA_real_),
         comments_per_function = case_when(
           n_functions != 0 & (!is.na(n_functions)) ~ n_example_comments/n_functions,
           TRUE ~ NA_real_
           ),
         comments_per_example = case_when(
           n_examples  != 0 ~ n_example_comments/n_examples ,
           TRUE ~ NA_real_
           ),
         has_examples = case_when(
           n_examples > 0 ~ TRUE,
           TRUE ~ FALSE
           )
         )
```

First, we'll carry out some analyses at the package level.

```{r package-analyses}
# difference in proportion of packages with vignettes ------------------
# find the point estimate
diff_prop_has_vignettes <- packages %>%
  mutate(has_vignettes = case_when(
    num_vignettes > 0 ~ TRUE,
    TRUE ~ FALSE
  )) %>%
  group_by(is_tidy) %>%
  summarize(mean_has_vignettes = mean(has_vignettes)) %>%
  select(mean_has_vignettes) %>%
  pull() %>%
  diff()


diff_prop_has_vignettes

# generate the null distribution and get the p-value
packages %>%
  mutate(has_vignettes = case_when(
    num_vignettes > 0 ~ TRUE,
    TRUE ~ FALSE)
    ) %>%
  specify(has_vignettes ~ is_tidy, success = "TRUE") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1e4, type = "permute") %>%
  calculate(stat = "diff in props", order = c("TRUE", "FALSE")) %>%
  get_p_value(obs_stat = diff_prop_has_vignettes, direction = "both")


# difference in proportion of packages with master help-files --------------
# find the point estimate
diff_prop_has_master_help <- packages %>%
  group_by(is_tidy) %>%
  summarize(mean_has_master_help = mean(has_master_help)) %>%
  select(mean_has_master_help) %>%
  pull() %>%
  diff()

diff_prop_has_master_help

# generate the null distribution and get the p-value
packages %>%
  specify(has_master_help ~ is_tidy, success = "TRUE") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1e4, type = "permute") %>%
  calculate(stat = "diff in props", order = c("TRUE", "FALSE")) %>%
  get_p_value(obs_stat = diff_prop_has_master_help, direction = "both")
```

Now, moving on to function-level analyses,

```{r function-level}
# difference in mean char count of function help-files --------------
# find the point estimate
diff_mean_n_char <- functions %>%
  group_by(is_tidy) %>%
  summarize(mean_n_char = mean(n_char, na.rm = TRUE)) %>%
  select(mean_n_char) %>%
  pull() %>%
  diff()

diff_mean_n_char

# generate the null distribution and get the p-value
functions %>%
  specify(n_char ~ is_tidy) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1e4, type = "permute") %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE")) %>%
  get_p_value(obs_stat = diff_mean_n_char, direction = "both")


# difference in mean number of examples in function help-files --------------
# find the point estimate
diff_mean_n_examples <- functions %>%
  group_by(is_tidy) %>%
  summarize(mean_n_examples = mean(n_examples, na.rm = TRUE)) %>%
  select(mean_n_examples) %>%
  pull() %>%
  diff()

diff_mean_n_examples

# generate the null distribution and get the p-value
functions %>%
  specify(n_examples ~ is_tidy) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1e4, type = "permute") %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE")) %>%
  get_p_value(obs_stat = diff_mean_n_examples, direction = "both")


# difference in mean number of comments in function help-file examples ----------
# find the point estimate
diff_mean_n_example_comments <- functions %>%
  group_by(is_tidy) %>%
  summarize(mean_n_example_comments = mean(n_example_comments, na.rm = TRUE)) %>%
  select(mean_n_example_comments) %>%
  pull() %>%
  diff()

diff_mean_n_example_comments

# generate the null distribution and get the p-value
functions %>%
  specify(n_example_comments ~ is_tidy) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1e4, type = "permute") %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE")) %>% 
  get_p_value(obs_stat = diff_mean_n_example_comments, direction = "both")

# difference in mean number of functions per help-file --------------
# find the point estimate
diff_mean_n_functions <- functions %>%
  group_by(is_tidy) %>%
  summarize(mean_n_functions = mean(n_functions, na.rm = TRUE)) %>%
  select(mean_n_functions) %>%
  pull() %>%
  diff()

diff_mean_n_functions

# generate the null distribution and get the p-value
functions %>%
  specify(n_functions ~ is_tidy) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1e4, type = "permute") %>%
  calculate(stat = "diff in means", order = c("TRUE", "FALSE")) %>% 
  get_p_value(obs_stat = diff_mean_n_functions, direction = "both")
```

To give context on the distribution of download numbers for packages in and out of the tidyverse,

```{r, warning = FALSE, message = FALSE}
# plot this difference
downloads_by_camp_viz <- pkgs_sum %>%
  mutate("Package Camp" = case_when(
    is_tidy ~ "Tidyverse",
    TRUE ~ "Non-Tidyverse"
  )) %>%
  ggplot() +
  aes(x = downloads_last_month, col = `Package Camp`) +
  geom_density() + 
  scale_x_log10() + 
  labs(x = "Log(Downloads Last Month)",
       y = "Relative Density") + 
  theme_minimal() +
  scale_color_brewer(type = "qual", palette = 1) +
  theme(text = element_text(family = "serif"))

downloads_by_camp_viz

# ...and save it!
ggsave(plot = downloads_by_camp_viz, 
       filename = "paper/figures/downloads_by_camp.png",
       width = 6, height = 4, units = "in")

# find observed difference (this is a population parameter)
pkgs_sum %>%
  group_by(is_tidy) %>%
  summarize(mean_num_dls = mean(downloads_last_month, na.rm = TRUE),
            n = n())
```


